The Controlled Choice Design and Paternalism in Pawnshop Borrowing

### 2023-10-21 Tables/Attrition.tex
cleaning of project folder - build replication files
stabilize cleaning files & internal validity analysis for number of pawns balance

% Table generated by Excel2LaTeX from sheet 'Attrition'
\begin{tabular}{lcccc}
\toprule
      &       & \multicolumn{3}{c}{Commitment arms} \\
\cmidrule{3-5}      & Control & Forced & Choice & p-value \\
\midrule
\midrule
Number of branch-day pawns & 34    & 34    & 36    & 0.52 \\
      & (3)   & (2.9) & (1.8) &  \\
\multicolumn{1}{p{12.82em}}{Ended up pawning} & 0.98  & 0.97  & 0.97  & 0.62 \\
      & (0.01) & (0.01) & (0.01) &  \\
Survey response rate & 0.79  & 0.76  & 0.77  & 0.62 \\
      & (0.02) & (0.02) & (0.01) &  \\
\midrule
Obs   & 1770  & 1954  & 2580  &  \\
\bottomrule
\bottomrule
\end{tabular}%

>>>>

% Table generated by Excel2LaTeX from sheet 'Attrition'
\begin{tabular}{lcccc}
\toprule
      &       & \multicolumn{3}{c}{Commitment arms} \\
\cmidrule{3-5}      & Control & Forced & Choice & p-value \\
\midrule
\midrule
Number of branch-day pawns & 31    & 31    & 36    & 0.19 \\
      & (2)   & (2.1) & (2)   &  \\
\multicolumn{1}{r}{median} & 27    & 28    & 31    & 0.15 \\
\midrule
Number of branch-day borrowers & 21    & 23    & 25    & 0.1 \\
      & (1.2) & (1.4) & (1.3) &  \\
\multicolumn{1}{r}{median} & 19    & 21    & 21    & 0.46 \\
\midrule
\multicolumn{1}{p{15.82em}}{Ended up pawning} & 0.98  & 0.97  & 0.97  & 0.62 \\
      & (0.01) & (0.01) & (0.01) &  \\
Survey response rate & 0.79  & 0.76  & 0.77  & 0.62 \\
      & (0.02) & (0.02) & (0.02) &  \\
\midrule
Obs   & 101   & 97    & 129   &  \\
\bottomrule
\bottomrule
\end{tabular}%

### 2023-10-21 donde.tex
stabilize cleaning files & internal validity analysis for number of pawns balance

The combination of voluntary and forced commitment within the RCT allows us to address several key questions. First, do structured repayment contracts lower financial costs for pawnshop borrowers? Second, do borrowers recognize this benefit, demanding commitment in sufficient numbers?  Finally, and most uniquely given the presence of a forced commitment arm, we are able to ask whether the \textit{right} borrowers voluntarily demand commitment. In particular, we use the choice, control, and forced arms to point identify the Treatment on the Treated (TOT) and the Treatment on the Untreated (TUT) effects, along with the average selection on gains (ASG) and average selection bias (ASB).

@@

Two other recent papers also shed some light on private paternalism. In the context of food choice, \cite{Sprenger} show that individuals with the most time-inconsistent preferences are actually least likely to demand commitment, suggesting a lack of awareness as to self-control problems. In the context of school choice, \cite{Walters} finds that students who select into more effective schools have smaller treatment effects from attending than those who do not.
Our paper differs from these in both its empirical setting and its methodology.
As far as we are aware, ours is the first experiment to combine control, forced treatment, and treatment choice arms and explain how such a design can be used to point identify TUT effects and a range of other interesting causal parameters under minimal assumptions.
While \cite{fowlie2021default} likewise exploited a three-armed experimental design in their study of the effect of electricity pricing, they identified two TOT effects for different groups of ``treated'' households, whereas we simultaneously identify the TOT and TUT effects defined with respect to a single ``treated'' group of borrowers.
We call our approach the ``Controlled Choice design'' for short.
Other approaches to going beyond local average treatment effects (LATE) from the literature either assume no unobserved selection-on-gains \citep{aronow2013beyond,angrist2013extrapolate} (the ``LATE-and-reweight'' approach) or rely on a combination of instruments with rich support and additional structural assumptions such as additive separability of observed and unobserved determinants of treatment effect heterogeneity \citep{heckman2007econometric, cornelissen2018benefits} (the marginal treatment effects approach).
\cite{Walters} takes the latter approach, combining a distance instrument with additional structural assumptions, obtaining model-based TUT and ATE estimates that differ substantially from those implied by a LATE-and-reweight approach. 
In contrast, we leverage our experimental design to point identify the TUT effect using a binary instrument without the need for a structural model, relying instead on relatively simple exclusion restrictions. This restriction has testable implications that we fail to reject. Rather than assuming away unobserved selection-on-gains, as in the LATE-and-reweight approach, we estimate it directly.
Unlike \cite{Sprenger} we directly identify the TUT, obviating the need to first elicit preferences before testing for negative selection.
In a returns-to-education setting, \cite{oreopoulos2006estimating}, uses a similar direct approach to identify a LATE \emph{nearly} equals a TUT effect: the 1944 Butler Act increased the share of British children who remained in school until age 15 from 43\% to over 90\%. Because 100\% of borrowers in our forced arm are treated, we identify the TUT rather than an approximation to it.

>>>>

The combination of voluntary and forced commitment within the RCT allows us to address several key questions. First, do structured repayment contracts lower financial costs for pawnshop borrowers? Second, do borrowers recognize this benefit, demanding commitment in sufficient numbers?  Finally, and most uniquely we are able to ask whether the \textit{right} borrowers voluntarily demand commitment. This ability comes from the unique three-armed experimental design, which we call the ``controlled choice design'' for short. As we show below, our design point identifies not only the overall effect of forced commitment--the average treatment effect (ATE)--but also the effect of commitment for borrowers who would \emph{voluntarily} choose it--the Treatment on the Treated (TOT) effect--along with the effect of commitment for borrowers who would \emph{not}--the Treatment on the Untreated (TUT) effect. Our design can be viewed as a juxtaposition of \emph{two} randomized encouragement designs, each with one-sided non-compliance; one of them identifies the TOT while the other identifies the TUT. By identifying both the TOT and TUT effects in the same experiment, the controlled choice design allows us to examine the empirical relevance of ``selection on gains'' also known as Roy-type selection into treatment: do borrowers who voluntarily choose commitment have higher average treatment effects than those who do not? The ``average selection on gains''--$\text{ASG} = \text{TOT} - \text{TUT}$ answers this question directly. The controlled choice design also point identifies the average selection bias (ASB)--the average difference in untreated potential outcomes for those who choose commitment relative to those who do not--along with the average selection on levels (ASL)--the analogous comparison for treated (commitment) potential outcomes. Taken together, these causal effects allow us to ``go under the hood'' of our baseline ATE results, and paint a more nuanced and economically-relevant picture of the effects of commitment. We are unaware of any other paper that simultaneously identifies all of these causal effects without recourse to additional structural modeling assumptions.

@@

Two other recent papers also shed some light on private paternalism. In the context of food choice, \cite{Sprenger} show that individuals with the most time-inconsistent preferences are actually least likely to demand commitment, suggesting a lack of awareness as to self-control problems. In the context of school choice, \cite{Walters} finds that students who select into more effective schools have smaller treatment effects from attending than those who do not.
Our paper differs from these in both its empirical setting and its methodology.
As far as we are aware, ours is the first experiment to combine control, forced treatment, and treatment choice arms and explain how such a design can be used to point identify TUT effects and a range of other interesting causal parameters under minimal assumptions.
While \cite{fowlie2021default} likewise exploited a three-armed experimental design in their study of the effect of electricity pricing, they identified two TOT effects for different groups of ``treated'' households, whereas we simultaneously identify the TOT and TUT effects defined with respect to a single ``treated'' group of borrowers.
%We call our approach the ``Controlled Choice design'' for short.
Other approaches to going beyond local average treatment effects (LATE) from the literature either assume no unobserved selection-on-gains \citep{aronow2013beyond,angrist2013extrapolate} (the ``LATE-and-reweight'' approach) or rely on a combination of instruments with rich support and additional structural assumptions such as additive separability of observed and unobserved determinants of treatment effect heterogeneity \citep{heckman2007econometric, cornelissen2018benefits} (the marginal treatment effects approach).
\cite{Walters} takes the latter approach, combining a distance instrument with additional structural assumptions, obtaining model-based TUT and ATE estimates that differ substantially from those implied by a LATE-and-reweight approach. 
In contrast, we leverage our experimental design to point identify the TUT effect using a binary instrument without the need for a structural model, relying instead on relatively simple exclusion restrictions. This restriction has testable implications that we fail to reject. Rather than assuming away unobserved selection-on-gains, as in the LATE-and-reweight approach, we estimate it directly.
Unlike \cite{Sprenger} we directly identify the TUT, obviating the need to first elicit preferences before testing for negative selection.
In a returns-to-education setting, \cite{oreopoulos2006estimating}, uses a similar direct approach to identify a LATE \emph{nearly} equals a TUT effect: the 1944 Butler Act increased the share of British children who remained in school until age 15 from 43\% to over 90\%. Because 100\% of borrowers in our forced arm are treated, we identify the TUT rather than an approximation to it.
